Refunctionalization of Abstract Abstract Machines (Functional Pearl)

### 2018-06-25 paper/paper.tex
more revisions

Defining a language by building an interpreter for it can be traced to the very
early days of programming languages research \cite{landin1966next, Reynolds:72}.
Nowadays, even an undergraduate student in computer science is able to build toy 
languages through interpreters.
But building a sound abstract interpreter remained an esoteric and difficult task 
until very recently.

@@

The CESK machine is an abstract machine for describing semantics of and evaluating 
$\lambda$-calculus \cite{felleisen1987calculus}. The CESK machine models program execution 
as state transitions in a small-step fashion. As its name suggests, a machine state has 
four components:
1) \textit{Control} is the expression currently being evaluated.
2) \textit{Environment} is a map that contains the address of a variable in the lexical scope.
3) \textit{Store} models the heap of a program as a map from addresses to values.
  The address space consists of numbers (0-indexed).
  In our toy language, the only category of value is a closure, i.e., a function paired with
  an environment.
4) \textit{Continuation} represents the program's call stack. In this paper, we instantiate the 
  stack as a list of frames because the ANF simplifies the evaluation context. For direct % TODO
  style $\lambda$-calculus, the continuations of CEK/CESK machines can be presented by
  variants of a data type.

The Scala representations for the components of the CESK machine are as follows:

@@

It is worth noting that the continuation class @Kont@ is defined as a list of
frames, where a frame can be considered a reduction context in reduction semantics.
We represent frames using the @Frame@ class, which stores the information of a single
call-site, i.e., the information that can be used to resume the interrupted computation.
A @Frame@ constitutes a variable name @x@ to be bound later, and a control component
to which the program may resume, and its environment.

@@

Function @continue@ is mainly what we have seen for the second case in @drive_step@,
except that @ndcontinue@ is called when the normal continuation is empty.
Since we are analyzing a language in ANF, the first-order data representation of
continuations is simplified to a Scala @List@. An empty list @Nil@ represents the %TODO
halt of a computation path, whereas a cons @::@ means the top frame of the list
holds the binding variable and context.
An analogy to this in standard CEK machines would be an abstract class (e.g., @cont@)
with three variants: @halt@, @arg(value, cont)@, and @arg(exp, env, cont)@.

>>>>

Defining a language by building an interpreter for it can be traced to the very
early days of programming languages research \cite{landin1966next, Reynolds:72}.
Nowadays, even an undergraduate student in computer science is able to make his own
programming language by building an interpreter.
But building a sound abstract interpreter remained an esoteric and difficult task 
until very recently.

@@

The CESK machine is an abstract machine for describing semantics of and evaluating 
$\lambda$-calculus \cite{felleisen1987calculus}. The CESK machine models program execution 
as state transitions in a small-step fashion. As its name suggests, a machine state has 
four components:
1) \textit{Control} is the expression currently being evaluated.
2) \textit{Environment} is a map that contains the address of a variable in the lexical scope.
3) \textit{Store} models the heap of a program as a map from addresses to values.
  The address space consists of numbers (0-indexed).
  In our toy language, the only category of value is a closure, i.e., a function paired with
  an environment.
4) \textit{Continuation} represents the program's call stack. In this paper, we instantiate the 
stack as a list of frames.

%For direct style $\lambda$-calculus, the continuations of CEK/CESK machines can be presented by variants of a data type.
The Scala representations for the components of the CESK machine are as follows:

@@

It is worth noting that the continuation class @Kont@ is defined as a list of
frames, where a frame can be considered a reduction context in reduction semantics.
An empty list represents empty call stack; otherwise, the head frame in the list
represents the innermost context, and the reduction result of this frame will be used
to fill the ``hole'' of its following element in the list.
We represent frames using the @Frame@ class, which stores the information of a single
call-site, i.e., the information that can be used to resume the interrupted computation.
A @Frame@ constitutes a variable name @x@ to be bound later, and a control component
to which the program may resume, and its environment.

@@

Function @continue@ is mainly what we have seen for the second case in @drive_step@,
except that @ndcontinue@ is called when the normal continuation is empty.
It works on the list representation of context: An empty list @Nil@ represents the
halt of a computation path, whereas a cons @::@ means the top frame of the list
holds the binding variable and context.
An analogy to this in standard CEK machines would be an abstract class (e.g., @cont@)
with three variants: @halt@, @arg(value, cont)@, and @arg(exp, env, cont)@.
